Truthfulness Verification System

### 2011-03-09 trunk/proposal/paper.tex
added progress section

\subsection{Truthfulness verification }
\subsubsection {Building queries from the data supplied by the user}

We take the content words from the topic unit($TU$), along with each of the
alternative units($AU$), and plugin the datatype($t$) to form a number of
queries.
\begin{align*}
q_{i} = \{{TU, AU_{i}, t_{AU_{i}}} \} \qquad \mbox {where $i = 1 \cdots n$} \\
\end{align*}
where $n$ is the number of alternative units generated by the first phase of
the existing system. By quering the various ontologies described above we will
get results which might have the following cases:

\begin{align*}
   freebase(q_{i}) &= \phi \\
    = r_{i}
\end{align*}

From what we have explored till now, by quering Freebase we can get the
various property, value pairs associated with an entity which is the $TU$ in
our case. In case the result is $\phi$, it suggests that we do not have
information for this particulat combination $TU$ and $AU_{i}$. Due to the richness of the information in Freebase we can assume with a considerable degree of confidence
that this combination is untrue. This is one way verification, but in case of
non-null results, how do we filter and rank the results with the available set
of information is something we can not formulate at this stage with out further
exploration.

\subsubsection{Application of Machine learning algorithms}
Though usage of machine learning algorithm seems to be very exciting approach at
this point, we are also skeptical about it. Supervised classifiers might not be
benficial for us due to the limited dataset for training and the potentialy
unlimited variety of inputs to classify. However, if we can use Graph
Based Partially Supervised Learning\cite{conf/acl/CelikyilmazTH09} or
Spreading Activation \cite{Cohen1987255} on the information extacted from the
rich ontologies is something we plan to explore.

>>>>

%\subsection{Truthfulness verification }
%\subsubsection {Building queries from the data supplied by the user}
%
%We take the content words from the topic unit($TU$), along with each of the
%alternative units($AU$), and plugin the datatype($t$) to form a number of
%queries.
%\begin{align*}
%q_{i} = \{{TU, AU_{i}, t_{AU_{i}}} \} \qquad \mbox {where $i = 1 \cdots n$} \\
%\end{align*}
%where $n$ is the number of alternative units generated by the first phase of
%the existing system. By quering the various ontologies described above we will
%get results which might have the following cases:
%
%\begin{align*}
%   freebase(q_{i}) &= \phi \\
%    = r_{i}
%\end{align*}
%
%From what we have explored till now, by quering Freebase we can get the
%various property, value pairs associated with an entity which is the $TU$ in
%our case. In case the result is $\phi$, it suggests that we do not have
%information for this particulat combination $TU$ and $AU_{i}$. Due to the richness of the information in Freebase we can assume with a considerable degree of confidence
%that this combination is untrue. This is one way verification, but in case of
%non-null results, how do we filter and rank the results with the available set
%of information is something we can not formulate at this stage with out further
%exploration.
%
%\subsubsection{Finding Sense of Words}
%To classify a statement as true or false we need to get the semantic of sense of the words of the sentences. For example consider the following sentence. 
%
%\emph{"Tom Hanks was the lead actress in the movie sleepless in Seattle"}
%
%Now our query results from ontologies will give us the fact that Tom hanks is a male actor. Now using wordnet we will be able to find that actress means female actor. So by simple inference our system will be able to classify the above example as untruthful.
%
%However, finding sense of words in a complex problem in general and not easy to solve for all cases all the time. For example, consider the following sentence.
%
%\emph{"Filipino is the primary language of Philipines"}
%
%now Wikipedia has "Filipino is the national language of Philipines". Both sentences have same meaning but it is not easy to relate primary with national because they are not each others synonym as separate words. They become synonym depending on the context they are used such as above example. At this point, we don't have any generic strategy to handle these kind of situations. However, we will explore more to find possible ways to solve these problems. And one possible way may be to use some natural language processing techniques. Specifically, we will look into NLTK library to solve these kind of issues.
%
%\subsubsection{Use of Co-reference Program}
%Coreference resolution is the process of determining whether two
%expressions in natural language refer to the same entity in the world.
%\cite{doi:10.1162/089120101753342653}.
%In the context of previous example we can use co-reference resolution
%if we have sentences like
%
%
%\emph{"Tom Hanks is the lead actor in the movie Sleepless in Seatle.
%He was nominated for the Golden Globe awards for this movie in 1994."
%}
%
%The pronoun "He" refers to Tom Hanks and would imply
%the document is about a male, allowing us to disambiguate Tom Hank's
%gender. The popular conference resolution api-s in use are those from
%LingPipe\cite{LingPipecoref} and OpenNLP \cite{OpenNLPcoref}. We have
%not yet evaluated them, and pondering if it would add sufficient value
%to the system as the topic units need not be so specific. More
%ontological information might already be available.
%
%
%
%
%\subsubsection{Application of Machine learning algorithms}
%Though usage of machine learning algorithm seems to be a good approach at
%this point. Consider the following example.
%
%\emph{"Tom Hanks was the lead actress in the movie sleepless in Seattle"}
%
%Now If the algorithm can learn that Tom hanks is a male and so he can't be an actress then our system should be able to classify the above statement as false. 
%
%However, we will need to explore more to be sure that we can apply machine learning algorithms successfully. Supervised classifiers might not be beneficial for us due to the limited dataset for training and the potentially unlimited variety of inputs to classify. However, if we can use Graph Based Partially Supervised Learning\cite{conf/acl/CelikyilmazTH09} or Spreading Activation \cite{Cohen1987255} on the information extracted from the rich ontologies is something we plan to explore.

\subsection{Step-by-step Algorithm for Truthfulness Verification}

\subsection{Examples of Truthfulness Verification}

\subsubsection{Use of Freebase}
Consider the following sentence:

\emph{"Tom Hanks was the lead actress in the movie Sleepless in Seattle"}

Content words for this sentence are "Tom Hanks", "lead", "actress", "movie", "Sleepless in Seattle".

Now for actress we will not find any result from Freebase query. So we will get synset of actress from wordnet. Wordnet synset for actress content for actress is \emph{"female actor"}

Now for the noun phrase \emph{"sleepless in Seattle"}, we can generate "id" for the query as "sleepless in seattle". But this id will be associated with many properties. To select the relevant property we can use synset obtained from the actress. And this synset has actor, which is one of the property for id "Sleepless in Seattle". Hence the Freebase query can be following:

\begin{lstlisting}

[{
  "id" : "/en/sleepless_in_seattle"
  "/film/film/starring" : [{ "actor" : null }]
}]

Now the result of this query returns following output:

{
  "code":          "/api/status/ok",
  "result": [{
    "/film/film/starring": [
      {
        "actor": "Tom Hanks"
      },
      {
        "actor": "Meg Ryan"
      },
      {
        "actor": "Bill Pullman"
      },
      {
        "actor": "Rosie O'Donnell"
      },
      {
        "actor": "Rob Reiner"
      },
      {
        "actor": "Victor Garber"
      },
      {
        "actor": "Gaby Hoffmann"
      },
      {
        "actor": "Carey Lowell"
      },
      {
        "actor": "David Hyde Pierce"
      },
      {
        "actor": "Ross Malinger"
      },
      {
        "actor": "Frances Conroy"
      },
      {
        "actor": "Rita Wilson"
      }
    ],
    "id": "/en/sleepless_in_seattle"
  }],
  "status":        "200 OK",
  "transaction_id": "cache;cache03.p01.sjc1:8101;2011-03-09T05:09:44Z;0032"
}

One of the actors in this result set is "Tom Hanks" that matches with our content word "Tom Hanks" in the given sentence. Now as we know earlier that actress means female actor, we can use the keyword female to find the fact that female is type of gender and Freebase id of "tom hanks" has a property gender associated with it. So we can formulate following query.

\begin{lstlisting}

[{
  "id" : "/en/tom_hanks"
  "/people/person/gender" : {}
}]

The result of this query is following:

{
  "code":          "/api/status/ok",
  "result": [{
    "/people/person/gender": {
      "id":   "/en/male",
      "name": "Male",
	.
	.
	.
}

Here the gender is Male, which contradicts with our gender female. Hence we can decide that this statement is false.

For finding the true statement i.e. the actress we can use all actors obtained in the first query result and form second query with their names and output the truthful sentence if we get female as the gender.

\subsubsection{Use of Wikipedia}
Although structured databases such as Freebase, DBPedia extract the information from Wikipedia and format them into a structured and query-friendly way, it does not contain all the relevant information all the time. For example, consider the following sentence:

\emph{Less Paul invented the electric guitar}

The wikipedia page "http://en.wikipedia.org/wiki/Electric\_guitar" contains the inventor of the electric guiter in one of the sentences in the body. But this information is not available in the Freebase.

We can use a very simple and elegant method to find the relevant information from Wikipedia by using simple web-scrapping and pattern matching using regular expression. The process is described below.

After getting the wikipedia page we can convert the html file into text file and then we can separate all the sentences in different lines. The sentence is the part of the whole text that exist between two full-stops. Another important observation is that the relevant information will be within the same sentence in the question-answer type sentences almost all the time. So we can match the content words to the sentences extracted from Wikipedia page and verify the truthfulness of the sentence.

For the example sentence above, after processing the Wikipedia page and matching by inventor we get following two sentences:

\emph{"Commercial production began in late summer of 1932 by Electro-Patent-Instrument Company Los\_Angeles, a partnership of Adolph\_Rickenbacker, PaulBarth and George\_Beauchamp, the inventor"}

\emph{In a morecontemporary style, Little Willie Joe, the inventor of the Unitar, had a rhythmand\_blues instrumental hit in the 1950s with "Twitchy", recorded with the ReneHall Orchestra}

Now we can match the doubt unit Les Paul with these two sentences and we will not have any match. Hence we can conclude that the statement is not true.

Moreover, we can also look for truthful statement by using those above extracted sentences. We can POS (Parts-of-Speech) tagger to get the noun phrases and create the list of possible persons that invented the electric guitar and match them with the five alternative sentences given by T-verifier.

\subsubsection{Use of answer.com}
We found that answer.com also contains truthful sentences for the doubt sentences in most of the cases. Hence using answer.com seems to be a good idea for the truthfulness verification. We can use similar strategies used for Wikipedia described in the section above.

\section{Progress up to now}
Following are the progress of us in these two weeks
\begin{itemize}
\item Formulated the procedure of truthfulness verification in a more structured and robust way compared to the last project proposal report
\item As a result of the procedure formulation, the sections Step-by-step Algorithm for Truthfulness and Verification Examples of Truthfulness Verification are written in a complete revised way compared to the proposal
\item Read the documentation of the Freebase. The language for querying Freebase is called MQL and it is very complex in learn and apply for big project such as Truthfulness Verification. The documentation itself is also several hundred pages. However, we think that we will be able to use Freebase for our project in a successful way.
\end{itemize}

\section{Conclusion}
It this paper, we described current system for truthfulness verification of statements found in the web called T-Verifier and described the possible improvements to the current system. We will primarily use authentic information sources such as Wikipedia and Ontologies extracted from Wikipedia such as FreeBase, DBPedia, and YOGA to found the truthfulness of doubtful statements.
